Environments and eval in R

### 2021-04-13 analysis/environments.Rmd
Environments

```{r envir_integer}
eval_calls_integer <- eval_calls_explicit %>% filter(envir_type == "integer")

@@

```{r}
caller_envs <- env_calls %>% filter(!is.na(hierarchy), is.na(specific_env) | specific_env == "_callee", is.na(new_env))

@@

```{r}

pack_envs <- env_classes %>% filter(is.na(new_env), is.na(hierarchy), specific_env != "global")
nb_pack_sites <- n_distinct(pack_envs$src_ref)

@@

synthetic_envs <- env_classes %>% filter(envir_type %in% c("list", "NULL") | !is.na(new_env) | specific_env == "empty")
nb_synthetic_sites <- n_distinct(synthetic_envs$src_ref)

@@

```{r}
pack_env_sites <- pack_envs %>% select(src_ref) %>% mutate(category = "package") 
function_env_sites <- bind_rows(select(caller_envs, src_ref), select(eval_calls_integer, src_ref)) %>% mutate(category = "function")
synthetic_env_sites <- synthetic_envs %>% select(src_ref) %>% mutate(category = "synthetic")
global_env_sites <- global_envs %>% select(src_ref)%>% mutate(category = "global")

category_per_site <- bind_rows(pack_env_sites, function_env_sites, synthetic_env_sites, global_env_sites) %>%
  group_by(src_ref) %>% summarise(n_categories = n_distinct(category))

@@

@@

# Specific envir with their environment class

For the default argument:

```{r}
eval_calls_default %>% count(environment_class, wt = duplicates, sort = TRUE) %>%
  mutate(pc = round(100 * n / sum(n), 2), cpc = cumsum(pc))

eval_calls_default %>% group_by(environment_class) %>% summarise(n = n_distinct(src_ref)) %>%
  arrange(desc(n)) %>%
   mutate(pc = round(100 * n / sum(n), 2), cpc = cumsum(pc))

# caller-2- is eval.parent
eval_calls_default %>% filter(environment_class == "caller-2-") %>% count(ev_variant)

`eval` is not problematic:

```{r}
eval_calls_default %>% filter(ev_variant == "eval") %>% count(environment_class, wt = duplicates, sort = TRUE)
```

But `eval.parent` is:

```{r}
eval_calls_default %>% filter(ev_variant == "eval.parent") %>% count(environment_class, wt = duplicates, sort = TRUE)
```

It might be because `parent.frame` is equivalent to  `sys.frame(sys.parent(n))` and `sys.parent(n)` can get surprising result of it is called near the top level...

For `parent.frame()` :

```{r}
E_envir %>% filter(envir == "parent.frame()") %>% count(environment_class, wt = duplicates, sort = TRUE) %>%
  mutate(pc = round(100 * n / sum(n), 2), cpc = cumsum(pc))

E_envir %>% filter(envir == "parent.frame()") %>% group_by(environment_class) %>%
  summarise(n = n_distinct(src_ref)) %>%
  arrange(desc(n)) %>%
   mutate(pc = round(100 * n / sum(n), 2), cpc = cumsum(pc))

For `new.env()`:

```{r}
E_envir %>% filter(envir == "new.env()") %>% count(environment_class, wt = duplicates, sort = TRUE) %>%
  mutate(pc = round(100 * n / sum(n), 2), cpc = cumsum(pc))

E_envir %>% filter(envir == "new.env()") %>% group_by(environment_class) %>%
  summarise(n = n_distinct(src_ref)) %>%
  arrange(desc(n)) %>%
   mutate(pc = round(100 * n / sum(n), 2), cpc = cumsum(pc))

For each envir expression, what is the most common environment class? (in calls, in sites)

```{r}
env_class_envir <- E_envir %>% group_by(envir) %>%
  group_by(environment_class, .add = TRUE) %>% summarise(n = n_distinct(src_ref)) %>%
  summarise(env_class = first(environment_class, order_by = order(n, decreasing = TRUE)), n = sum(n)) %>% 
  arrange(desc(n))
```

>>>>

```{r envir_integer}
eval_calls_integer <- eval_calls_explicit %>% filter(envir_type %in% c("integer", "double")) 
# we assume that the doubles are properly casted to integers here

@@

```{r}
caller_envs <- env_calls %>% filter(!is.na(hierarchy), is.na(new_env))
# Will count also caller-n-specific env

@@

```{r}
# 4 srcrefs will be counted twice, in package environments, and in caller environments
pack_envs <- env_classes %>% filter(is.na(new_env),  specific_env != "global")
nb_pack_sites <- n_distinct(pack_envs$src_ref)

@@

synthetic_envs <- env_classes %>% filter(envir_type %in% c("list", "NULL", "pairlist") | !is.na(new_env) | specific_env == "empty")
nb_synthetic_sites <- n_distinct(synthetic_envs$src_ref)

@@

```{r}
pack_env_sites <- pack_envs %>% select(src_ref, duplicates) %>% mutate(category = "package") 
function_env_sites <- bind_rows(select(caller_envs, src_ref, duplicates), select(eval_calls_integer, src_ref, duplicates)) %>% mutate(category = "function")
synthetic_env_sites <- synthetic_envs %>% select(src_ref, duplicates) %>% mutate(category = "synthetic")
global_env_sites <- global_envs %>% select(src_ref,duplicates)%>% mutate(category = "global")

env_categories <- bind_rows(pack_env_sites, function_env_sites, synthetic_env_sites, global_env_sites)

category_per_site <- env_categories %>%
  group_by(src_ref) %>% summarise(n_categories = n_distinct(category))

@@

#TODO: create tags for the poly categories df

@@

And what if we classify by using the most present category for site?
```{r}
majority_category <- env_categories %>% group_by(src_ref, category) %>% summarise(n_calls = sum(duplicates)) %>%
  summarize(category = first(category, order_by = order(n_calls, decreasing = TRUE)), n = n(), prop_first = max(n_calls) / sum(n_calls))

majority_category %>% group_by(category) %>% summarise(n = n_distinct(src_ref)) %>% arrange(desc(n)) %>% mutate(pc = ratio(n, nb_call_sites), cpc = cumsum(pc))
```
